### 2018-01-02 preamble.tex

\usepackage{xcolor}
\usepackage{url}

%\usepackage{endfloat}
%\usepackage{soul}

%% \usepackage{type1ec}
 \usepackage{subcaption}
% \usepackage{fullpage}
% \usepackage{enumerate}
% %\usepackage{graphicx}
% %\usepackage{graphics}
% \usepackage{multirow}
% \usepackage{comment}
% \usepackage{amsmath,amssymb,amsfonts,amsthm}
% \usepackage{bbm}
% \usepackage{setspace}
% \usepackage{verbatim}
% %\usepackage{natbib}
% \usepackage{bm}
% \usepackage{pdflscape}
% %\usepackage[dvipsnames]{xcolor}
% \usepackage{tikz}
% \usepackage{xr}
% %\usepackage{mathabx}
% %\usepackage{filecontents}
% %\usepackage{bibentry}
% %\usepackage{hanging}
% %\usepackage{apacite}
%\usepackage{hyperref}
\usepackage{makecell}
\usepackage{amsmath}
%\usepackage{endnotes}

%\let\footnote=\endnote

\title{Mastery Learning in Practice:\\
A (Mostly) Descriptive Analysis of Log Data from the Cognitive Tutor Algebra I Effectiveness Trial}

\author{{\large Anita Israni}\\University of Texas College of
  Education\\aisrani@gmail.com \and {\large Adam C Sales}\\University of Texas College of
  Education\\asales@utexas.edu  \and {\large John F Pane}\\RAND Corporation\\jpane@rand.org}

\date{}